---
editor_options:
  chunk_output_type: console
output:
  pdf_document: default
  html_document: default
---
Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
# library(dplyr)
# pf.fc_by_age_gender <- pf%>%
#   filter(!is.na(gender))%>%
#   group_by(age, gender)%>%
#   summarise(mean_friend_count = mean(friend_count),
#             median_friend_count = median(friend_count),
#             n = n())%>%
#   arrange(age, gender)
# 
# head(pf.fc_by_age_gender)

```
####The Plot that we created before:
```{r}
library(ggplot2)
# ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
#   geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```
####Now, we use a different way using the new data frame we just created to plot the identical picture.
```{r}
# ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
#   geom_line(aes(color = gender))
```


***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:
```{r}
# install.packages('tidyr')
# library('tidyr')
# spread(subset(pf.fc_by_age_gender, 
#        select = c('gender', 'age', 'median_friend_count')), 
#        gender, median_friend_count)
```

***

### Reshaping Data
Notes:

```{r}
# library('dplyr')
# library('tidyr')
# pf.fc_by_age_gender.wide <-
#   subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
#          !is.na(gender)) %>%
#   spread(gender, median_friend_count) %>%
#   mutate(ratio = male / female)

# head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# library('ggplot2')
# ggplot(aes(x = age, y = female / male), data = pf.fc_by_age_gender.wide) +
#   geom_line()
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# colnames(pf)
# pf$year_joined <- floor(2014 - pf$tenure / 365)
# head(pf)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
# pf$year_joined.bukcket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
# summary(pf$year_joined.bukcket)
# table(pf$year_joined.bukcket)
# ```





### Plot the Grand Mean

# ``` {r}
# ggplot(aes(x = age, y = friend_count),
#        data = subset(pf, !is.na(year_joined.bukcket))) +
#   geom_line(aes(color = year_joined.bukcket), stat = 'summary', fun.y = mean)+
#   geom_line(stat = 'summary', fun.y = mean, linetype = 2)
#```

***

### Friending Rate
Notes:

```{r Friending Rate}
# with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
# colnames(pf)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

# ggplot(aes(x = tenure, y = friendships_initiated / tenure),
#        data = subset(pf, tenure >= 1)) +
#   geom_line(aes(color = year_joined.bucket),
#             stat = 'summary',
#             fun.y = mean)
# 
# ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
#        data = subset(pf, tenure > 0)) +
#   geom_line(aes(color = year_joined.bucket),
#             stat = "summary",
#             fun.y = mean)
# 
# ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
#        data = subset(pf, tenure > 0)) +
#   geom_line(aes(color = year_joined.bucket),
#             stat = "summary",
#             fun.y = mean)
# 
# ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
#        data = subset(pf, tenure > 0)) +
#   geom_line(aes(color = year_joined.bucket),
#             stat = "summary",
#             fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:
```{r}
# yo <- read.csv('yogurt.csv')
# str(yo)
# #Change the id from an int to a factor
# yo$id <- factor(yo$id)
# str(yo)
```

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
# ggplot(aes(x = price), data = yo) +
#   geom_histogram(fill = '#F79420')
```
###To see how many distinct price we have in the dataset 'yo':
```{r}
# unique(yo$price)
# length(unique(yo$price))
# summary(yo$price)
# table(yo$price)
```

***

### Number of Purchases
Notes:
To see how many purchases in total each household has
```{r Number of Purchases}
# colnames(yo)
# yo <- mutate(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
# head(yo)
# 
# ggplot(aes(x = all.purchases), data = yo) +
#   geom_histogram(binwidth = 1)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
# ggplot(aes(x = time, y = price), data = yo) +
#   geom_jitter(alpha = 1/4, shape = 21, fill = '#F79420')

```
We can see that the price is increasing over time, we also see some lower price point scattered about the graph, this may be due to puchasers using coupons or something like that.
***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
# set.seed(4230)
# sample.ids <- sample(levels(yo$id), 16)
# 
# ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) +
#   geom_line() + facet_wrap(~id) + 
#   geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
# library(GGally)
# theme_set(theme_minimal(20))
# #Set the seed for reproducible results
# set.seed(1836)
# pf_subset <- pf[ ,c('age', 'dob_year','dob_month','gender','tenure')]
# ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
# nci <- read.table("nci.tsv")
# colnames(nci) <- c(1:64)
```

```{r}
# nci.long.samp <- melt(as.matrix(nci[1:200,]))
# names(nci.long.samp) <- c("gene", "case", "value")
# head(nci.long.samp)
# 
# ggplot(aes(y = gene, x = case, fill = value),
#   data = nci.long.samp) +
#   geom_tile() +
#   scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


